\documentclass{denniscv}

\begin{document}

\name{Dennis Chen}

\sociallist{
    \social{\faGlobe}{\href{https://dennisc.net}{dennisc.net}}\sep%
    \social{\Letter}{\href{mailto:dchen@dennisc.net}{dchen@dennisc.net}}\sep%
    \social{\faGithub}{\href{https://github.com/chennisden}{chennisden}}%
}
\cvQuote{Improving yourself daily means increasing your self-worth. Turn adversity into strength!}{Yugami}

\cat{Education}

\act{Amador Valley High School}{Fall 2019}{present}

\desc{Graduating in 2023. Current unweighted GPA is 4.0, weighted GPA is 4.4.}

\act{Linear Algebra}{Fall 2020}{}

\desc{A in the class. Dual enrolled in Chabot Community College during my sophomore year.}

\cat{Work}

\act{AlphaStar Academy}{December 2020}{July 2021}

\desc{Developed math curriculum as a part-time employee.}

\cat{Volunteering and Programs}

\act{Math Advance!}{2021}{present}

\desc{Current director and founder of math enrichment association.
We hold classes and competitions, most notably MAST and the MAT.
More details can be found on \href{https://www.mathadvance.org}{mathadvance.org}.}
% replace "math enrichment association" with "501(c)(3) mathematics non-profit"
% once the process is finished

\cat{Awards}

\act{USA Junior Math Olympiad (USAJMO) qualifier}{2020}{}

\act{American Invitational Mathematics Examination qualifier}{2019, 2020, 2021}{}

\cat{Publications}

\act{Proofs in Competition Math: Volume 1}{}{}

\refer{Toller, Alex, et al. Proofs in Competition Math. Vol. 1, Amazon Publishing, 2019.}

\act{Proofs in Competition Math: Volume 2}{}{}

\refer{Toller, Alex, et al. Proofs in Competition Math. Vol. 2, Amazon Publishing, 2019.}

\cat{Projects}

\act{bounce.cls}{}{}

\desc{Open-source \LaTeX{} class for articles and books.
GitHub repository can be found at
\href{https://www.github.com/chennisden/bounce}{github.com/chennisden/bounce}.}

\cat{Skills}

\act{HTML, CSS, Javascript, Java, \LaTeX, Asymptote, TikZ}{}{}

\desc{Experience developing web applications with React JS.
Worked with LaTeX formatting and logic.}

\end{document}